State of the Practice for Lattice Boltzmann Method Software

### 2022-01-03 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Edits to LBM SOP

\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{arydshln}
\usepackage{paralist}
\usepackage{multicol}

@@

    We analyze the state of software development practice for Lattice Boltzmann
    solvers by quantitatively and qualitatively measuring and comparing 24
    software packages for 10 software qualities (installability, correctness/
    verifiability, reliability, robustness, usability, maintainability,
    reusability, understandability, visibility/transparency and
    reproducibility). Our reproducible analysis method employs a measurement
    template (containing 103 measures that are manually and automatically
    extracted from the project repository) and developer interviews (consisting
    of a set of 20 questions). From the measurement results, we ranked the
    software using the Analytic Hierarchy Process (AHP), a multi-criteria
    decision making method appropriate for cases with a mix of qualitative and
    quantitative factors. Our ranking was roughly consistent with ranking the
    repositories by the number of GitHub stars, although the number one project
    by star count (Sailfish) is ranked 9th by our methodology. Our top three
    packages were ESPResSo, Ludwig, and Palabos (all in the top 10 by star
    count). We interviewed 4 developers to gain insight into their current pain
    points. Identified challenges include lack of development time, lack of
    funding, and difficulty with ensuring correctness.  Based on our assessment,
    we make the following recommendations for current and future LBM projects:
    provide a detailed user manual and tutorials, explicitly state the limits of
    the software, use user-friendly programming languages, consider a peer
    review process, communicate development and contribution information, and
    use continuous integration and project management tools.

@@

\begin{table}
\begin{center}
	\begin{tabular}{ p{4cm}p{3cm}p{3cm} }
		\hline
		Name & Released & Updated\\
		\hline
		DL\_MESO (LBE) & unclear & 2020 Mar\\
		ESPResSo & 2010 Nov & 2020 Jun\\
		ESPResSo++ & 2011 Feb & 2020 Apr\\
		lbmpy& unclear  & 2020 Jun  \\
		lettuce & 2019 May & 2020 Jul\\
		Ludwig& 2018 Aug & 2020 Jul\\
		LUMA& 2016 Nov   & 2020 Feb\\
		MechSys & 2008 Jun    & 2021 Oct\\
		Musubi & 2013 Sep & 2019 Aug\\
		OpenLB & 2007 Jul & 2019 Oct\\
		Palabos & unclear & 2020 Jul\\
		pyLBM & 2015 Jun&   2020 Jun\\
		Sailfish & 2012 Nov & 2019 Jun\\
		TCLB & 2013 Jun  & 2020 Apr\\
		waLBerla & 2008 Aug & 2020 Jul\\
		\hline
	\end{tabular}
	\caption{Alive Software Packages} \label{alivepackages}
\end{center}
\end{table}

@@

\begin{table}
	\begin{center}
		\begin{tabular}{ p{3cm}llllllll}
			\hline
			Name & Dim & Pll & Com & Rflx & MFl & Turb & CGE & OS\\
			\hline
			DL\_MESO (LBE) & 2, 3 & MPI/OMP & Y & Y & Y & Y & Y & W, M, L\\
			ESPResSo & 1, 2, 3 & CUDA/MPI & Y & Y & Y & Y & Y & M, L\\
			ESPResSo++ & 1, 2, 3 & MPI & Y & Y & Y & Y & Y & L\\
			HemeLB & 3 & MPI & Y & Y & Y & Y & Y & L\\
			laboetie & 2, 3 & MPI & Y & Y & Y & Y & Y & L\\
			LatBo.jl & 2, 3 & - & Y & Y & Y & N & Y & L\\
			LB2D-Prime & 2 & MPI & Y & Y & Y & Y & Y & W, L\\
			LB3D & 3 & MPI & N & Y & Y & Y & Y & L\\
			LB3D-Prime & 3 & MPI & Y & Y & Y & Y & Y & W, L\\
			lbmpy & 2, 3 & CUDA & Y & Y & Y & Y & Y & L\\
			lettuce & 2, 3 & CUDA & Y & Y & Y & Y & Y & W, M, L\\
			LIMBES & 2 & OMP & Y & Y & N & N & Y & L\\
			Ludwig & 2, 3 & MPI & Y & Y & Y & Y & Y & L\\
			LUMA & 2, 3 & MPI & Y & Y & Y & Y & Y & W, M, L\\
			MechSys & 2, 3 & - & Y & Y & Y & Y & Y & L\\
			MP-LABS & 2, 3 & MPI/OMP & N & Y & Y & N & N & L\\
			Musubi & 2, 3 & MPI & Y & Y & Y & Y & Y & W, L\\
			OpenLB & 1, 2, 3 & MPI & Y & Y & Y & Y & Y & W, M, L\\
			Palabos & 2, 3 & MPI & Y & Y & Y & Y & Y & W, L\\
			pyLBM & 1, 2, 3 & MPI & Y & Y & N & Y & Y & W, M, L\\
			Sailfish & 2, 3 & CUDA & Y & Y & Y & Y & Y & M, L\\
			SunlightLB & 3 & - & Y & Y & N & N & Y & L\\
			TCLB & 2, 3 & CUDA/MPI & Y & Y & Y & Y & Y & L\\
			waLBerla & 2, 3 & MPI & Y & Y & Y & Y & Y & L\\
			\hline
		\end{tabular}
		\caption{Features of Software Packages (Dim for Dimension (1, 2, 3),
			Pll for Parallel (CUDA, MPI, OpenMP (OMP)), Com for Compressible (Yes or No), Rflx for
			Reflexive Boundary Condition (Yes or No), MFl for Multi-fluid (Yes or
			No), Turb for Turbulent (Yes or No), CGE for Complex Geometries (Yes or No), OS for Operating System (Windows (W), macOS (M), Linux (L)))} \label{tbl_features}
	\end{center}
\end{table}

@@

All of the 24 software packages that were tested have installation instructions.
As noted previously, many of the 22 software packages that were part of the
original long list of 46 packages were removed due to not including
documentation or installation instructions. Of the 24 software packages that
were tested, most had installation instructions located in one place, often in
an instruction manual or on a web-page. Sometimes, like with Ludwig, incomplete
installation instructions are found on a home page, with more detailed
instructions located on another web-page, or within the documentation.
Maintainability and correctness of these instructions could be improved if all
the instructions were in one location.

All 24 packages on the short list can be installed on some Unix-like systems.
Eight packages could be installed on Windows, and five on macOS. Operating system requirements are listed in the documentation of 20 software packages, but compatible operating system versions are only specified in four packages (ESPResSo, LUMA, Mechsys, Sailfish). All but one of the software packages (TCLB) were tested on Ubuntu for this state of the practice assessment. TCLB was tested on CentOS, since this operating system is mentioned in its installation instructions.

@@

Seventeen software packages require less than 10 dependencies to be installed. All but one software package (LatBo.jl) require less than 20 dependencies. Some
packages may automatically install additional dependencies in the background.
Twenty of the software packages do not explicitly indicate software dependency
versions. Some software package installation issues, specifically those
occurring when manual installation of dependencies is required, may be avoided
if versions of dependencies are specified. Sixteen software packages do not have
detailed instructions for installing dependencies. Sixteen software packages
have less than 10 manual installation steps. If dependencies are installed in
one command then none of the software packages take more than 20 steps to
install. The average number of steps is about eight, and the fewest is two
(LB3D-Prime).

@@

Figure~\ref{Fig_Installability} shows the installability ranking of the software
packages using AHP. Software packages with a higher score (ESPResSo, MechSys, OpenLB, DL\_MESO, Palabos) tend to have one set of linear installation instructions that are written as if the person doing the installation has none of the dependencies installed. The instructions often list compatible operating system versions and include instructions for the installation of dependencies. The top ranked packages often incorporate some sort of automation of the installation process and have fewer manual installation steps. The number of dependencies a package has does not correlate with a higher score. The ability to validate the installation process, often through tutorials or test examples that include expected output, is correlated with a higher score. Furthermore, the top seven ranked packages are noted as being alive.

@@

Seventeen of the software packages include a requirements specification artifact
or explicitly reference domain theory, often only the latter. Software packages
that distribute requirements specification information, such as DL\_MESO (LBE),
generally keep it brief and include it within other documentation. This artifact
is often found within a user manual, on a web-page, or is mentioned in related
publications. In the latter case the user may need to spend significant time to
find this information.

Document generation tools are explicitly used by 13 software packages. Sphinx is
used by eight of them, and Doxygen is used by eight. Several of the packages use
both.

Tutorials are available for 19 of the software packages. Generally they are
linearly written and easy to follow. However, only eight tutorials provide an
expected output. It is not possible to verify the correctness of the output of
the software packages that are missing this key information. In these cases the
user may need to assume correctness if there are no visible errors.

Unit tests are only explicitly available for two of the software packages,
Ludwig and Musubi. Code modularization of most packages allow for users to create tests with varying degrees of effort. These tests allow developers and users to verify the correctness of fragments of the source code, and in doing so better assess the correctness of the entire package.

The use of continuous integration tools and techniques alludes to a more refined
development process where faults are isolated and better recognized. Only three of the packages (ESPResSo, Ludwig, Musubi) mentioned applying the practice of continuous integration in their development process.

@@

Figure~\ref{Fig_CorrectnessVerifiability} shows the surface correctness and
verifiability ranking of the software packages using AHP. Software packages with
a higher score tend to have a visible requirements specification or references
to theory documentation. They also explicitly use at least one document
generation tool that builds confidence of correctness. The top ranked software
packages all include an easy to follow getting started tutorial, and most of
these include expected output. The two top ranked packages, Ludwig and Musubi, indicate the use of unit testing in their documentation. They also incorporated continuous integration in the development process. Furthermore, eight of the top 10 ranked packages are noted as being alive.

@@

Of the 14 software packages that installed correctly and also have tutorials,
four (pyLBM, ESPResSo++, LIMBES, Ludwig) broke during tutorial testing. All of
these instances resulted in an error message being displayed. One error (pyLBM)
was due to a missing tutorial dependency, another (Ludwig) was due to an invalid
command despite following the tutorial, and the final two errors were vague
execution errors. Of the four broken tutorial instances, only the one that was
missing a dependency was recoverable.

@@

Figure~\ref{Fig_Reliability} shows the surface reliability ranking of the
software packages using AHP. Software packages with a high score either did not
break during installation, or the broken installation was recoverable. All of
the top five ranked packages have tutorials. The package pyLBM broke
during tutorial testing, but a descriptive error message helped in recovery.
Nine of the top 10 ranked packages are noted as being alive.

@@

Although the software packages all apply LBM to solve scientific computing
problems, the packages focus on varied CFD problems, with varying parameters,
and are technically different from each other (as shown in
Table~\ref{tbl_features}). Due to this, a comparison of performance is not
appropriate. In this project we instead looked through each software package's
artifacts for evidence that performance was considered. The artifacts of 18
software packages mentioned parallelization. This included GPU processing and
the CUDA parallel computing platform, which were mentioned in the artifacts of 6
packages (ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB). GPUs provide
superior processing power and speed compared to CPUs, and are often used for
scientific computing when a large amount of data is involved. The software
package TCLB is implemented in a highly efficient multi-GPU code to achieve
performance suitable for model optimization \citep{rutkowski2020open}. In the
Ludwig package, a so-called mixed mode approach is used where fine-grained
parallelism is implemented on the GPU, and MPI is used for even larger scale
parallelism \citep{gray2013ludwig}. While one software package (Sailfish)
required CUDA and GPU processing, some (ESPResSo, lbmpy, lettuce, pyLBM, TCLB)
have the option of using either the GPU or the CPU. The packages that require
GPU and CUDA have better performance at the expense of installability and
surface reliability.

@@

Software package artifacts were reviewed for the presence of a tutorial, a user
manual, documented user characteristics, and a user support model. In total 19
software packages have a tutorial, 13 have a user manual, and 11 have both. The
tutorials vary in scope and substance, and eight include expected output.
Most user manuals are in the form of a file that can be downloaded, while some
are rendered on a web-page. Some packages (waLBerla) do not have a user manual,
but do have useful documentation distributed on their web-pages. Expected user
characteristics are documented in five software packages (laboetie, LIMBES,
Ludwig, Musubi, Palabos). Users are typically scientists or engineers. Their background is often physics, chemistry, biophysics, or mathematics. All but one of the packages (LIMBES) have a user support model, and many of them have multiple avenues of user support. The most popular avenue of support is Git, followed by email and forums. One software package (OpenLB) has an FAQ page.

@@

Issue tracking is used in 23 software packages, 15 of which use Git, seven use
email, and one (SunlightLB) uses SourceForge. Most software packages that use
Git have most of their issues closed, and only three (laboetie, lettuce,
Sailfish) have less than 50 percent of their issues closed. All of the top five
overall ranked packages have most of their issues closed. Alive packages (11 use Git issue tracking) have 64\% of their issues closed, while dead packages (3 use Git issue tracking) have 71\% of their issues closed. This information is presented in Table~\ref{gitrepodata}. Furthermore, 13 packages that use Git for issue tracking use GitHub as a version control system, while two (Palabos, waLBerla) use GitLab. Of the other packages, one package (SunlightLB) uses CVS for issue tracking, and seven packages do not appear to use any issue tracking system.

\begin{table}
	\begin{center}
		\begin{tabular}{ p{3.5cm}p{3.5cm}p{3.5cm}p{2.5cm} }
			\hline
			Name & $\%$ Issues Closed & $\%$ Code Comments & Status\\
			\hline
			DL\_MESO (LBE) & Not Git & 8.06& Alive\\
			ESPResSo & 89.26 & 21.78 & Alive\\
			ESPResSo++ & 66.28 & 17.10 & Alive\\
			HemeLB & No Issues & 16.68 & Dead\\
			laboetie & 18.75 & 2.47 & Dead\\		
			LatBo.jl & 93.33 & 0.40 & Dead\\
			LB2D-Prime & Not Git & 13.61 & Dead\\
			LB3D & Not Git & 13.76 & Dead\\
			LB3D-Prime & Not Git & 14.34 & Dead\\
			lbmpy& 58.33  & 2.03 & Alive\\
			lettuce & 33.33 & 8.19 & Alive\\
			LIMBES & Not Git & 17.39 & Dead\\
			Ludwig& 60.00 & 20.70 & Alive\\
			LUMA& 85.71   & 0.20 & Alive\\
			MechSys & Not Git & 15.11 & Alive\\
			MP-LABS & 100.00 & 26.67 & Dead\\
			Musubi & Not Git & 24.19 & Alive\\
			OpenLB & Not Git & 22.43 & Alive\\
			Palabos & 89.47 & 17.76 & Alive\\
			pyLBM & 66.67& 16.12 & Alive\\
			Sailfish & 22.22 & 9.26 & Alive\\
			SunlightLB & Not Git & 17.67 & Dead\\
			TCLB & 60.32 & 6.02 & Alive\\
			waLBerla & 72.90 & 22.62 & Alive\\
			\hline
		\end{tabular}
		\caption{Git Repository Data} \label{gitrepodata}
	\end{center}
	\end{table}

Software package code files were further measured for the percentage of code
that is comments. The findings are presented in Table~\ref{gitrepodata}.
Packages with a higher percentage of comments were designated as more
maintainable. Comments represent more than 10 percent of code files in 16
packages, and the average percentage of code comments is about 14 percent. All
of the top five overall ranked packages have more than the average. LUMA has only 0.2 percent comments, the fewest of any package. This package has the most lines of source code, with over four million. The next largest package is ESPResSo++ with one million.

@@

Figure~\ref{Fig_Maintainability} shows the maintainability ranking of the
software packages using AHP. Software packages with a high score provide version
numbers on documents and source code releases, have an abundance of high quality
artifacts, and use an issue tracking tool and version control system. These
packages also appear to reasonably handle issue tracking, having most of their
issues closed. Their code files are well commented with more than 10 percent of
the code being comments. Furthermore, four of the top 5 ranked packages are
noted as being alive. MP-LABS is noted as being dead.\\

@@

\begin{table}
	\begin{center}
		\begin{tabular}{ p{3.5cm}p{2cm}p{2.5cm}p{2cm}p{2.5cm} }
			\hline
			Name & Text Files & Binary Files & LOC & Avg. LOC / Text File\\
			\hline
			DL\_MESO (LBE) & 310 & 51 & 170223 & 549\\
			ESPResSo & 1390& 83 & 195083 & 140\\
			ESPResSo++ & 1406& 30 & 165194 & 118\\
			HemeLB & 1102& 44 & 123806 & 112\\
			laboetie & 133& 1 & 48403 & 364\\		
			LatBo.jl & 41& 0 & 42172 & 1029\\
			LB2D-Prime & 82& 19 & 54755 & 668\\
			LB3D & 99 & 76 & 39766 & 402\\
			LB3D-Prime & 23& 6 & 12944 & 563\\
			lbmpy& 250 & 29 & 61632 & 247\\
			lettuce & 73 & 1 & 7660 & 105\\
			LIMBES & 26 & 1 & 4872 & 187\\
			Ludwig & 954 & 32 & 162518 & 170\\
			LUMA & 314 & 19 & 4399723 & 14011\\
			MechSys & 333 & 3 & 95707 & 287\\
			MP-LABS & 307 & 3 & 43124 & 140\\
			Musubi & 1347 & 1839 & 281879 & 209\\
			OpenLB & 1438 & 7 & 218406 & 152\\
			Palabos & 1974 & 71 & 563841 & 286\\
			pyLBM & 272 & 108 & 37234 & 137\\
			Sailfish & 632 & 11 & 69398 & 110\\
			SunlightLB & 36 & 1 & 7646 & 212\\
			TCLB & 594 & 7 & 49156 & 83\\
			waLBerla & 2643 & 69 & 873988 & 331\\
			\hline
		\end{tabular}
		\caption{Module Data} \label{moduledata}
	\end{center}
\end{table}

@@

Figure~\ref{Fig_Understandability} shows the surface understandability ranking
of the software packages using AHP. Software packages with a high score have a
consistent indentation and formatting style, and consistent, distinctive, and
meaningful code identifiers. They also have symbolic constants, and explicitly
identify mathematical and LBM algorithms. Their comments are clear and indicate
what is being done in the source code. The source code is well modularized and
structured. Furthermore, all of the top five ranked packages are noted as being
alive.

@@

\subsection{Pain Points} \label{painpoints}

>>>>

\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{arydshln}
\usepackage{paralist}
\usepackage{multicol}
\usepackage{booktabs}

@@

    We analyze the state of software development practice for Lattice Boltzmann
    solvers by quantitatively and qualitatively measuring and comparing 24
    software packages for 10 software qualities (installability, correctness/
    verifiability, reliability, robustness, usability, maintainability,
    reusability, understandability, visibility/transparency and
    reproducibility). Our reproducible analysis method employs a measurement
    template (containing 103 measures that are manually and automatically
    extracted from the project repository) and developer interviews (consisting
    of a set of 20 questions). From the measurement results, we ranked the
    software using the Analytic Hierarchy Process (AHP), a multi-criteria
    decision making method appropriate for cases with a mix of qualitative and
    quantitative factors. Our ranking was roughly consistent with ranking the
    repositories by the number of GitHub stars, although the number one project
    by star count (Sailfish) is ranked 9th by our methodology. Our top three
    packages were Ludwig, ESPResSo, and Palabos (all in the top 10 by star
    count). We interviewed 4 developers to gain insight into their current pain
    points. Identified challenges include lack of development time, lack of
    funding, and difficulty with ensuring correctness.  Based on our assessment,
    we make the following recommendations for current and future LBM projects:
    provide a detailed user manual and tutorials, explicitly state the limits of
    the software, use user-friendly programming languages, consider a peer
    review process, communicate development and contribution information, and
    use continuous integration and project management tools.

@@

\begin{table}
\begin{center}
	\begin{tabular}{ p{4cm}p{3cm}p{3cm} }
		\hline
		Name & Released & Updated\\
		\hline
		DL\_MESO (LBE) & unclear & 2020 Mar\\
		ESPResSo & 2010 Nov & 2020 Jun\\
		ESPResSo++ & 2011 Feb & 2020 Apr\\
		lbmpy& unknown  & 2020 Jun  \\
		lettuce & 2019 May & 2020 Jul\\
		Ludwig& 2018 Aug & 2020 Jul\\
		LUMA& 2016 Nov   & 2020 Feb\\
		MechSys & 2008 Jun    & 2020 Jul\\
		OpenLB & 2007 Jul & 2019 Oct\\
		Palabos & unclear & 2020 Jul\\
		pyLBM & 2015 Jun&   2020 Jun\\
		Sailfish & 2012 Nov & 2019 Jun\\
		TCLB & 2013 Jun  & 2020 Apr\\
		waLBerla & 2008 Aug & 2020 Jul\\
		\hline
	\end{tabular}
	\caption{Alive Software Packages} \label{alivepackages}
\end{center}
\end{table}

@@

\begin{table}
	\begin{center}
		\begin{tabular}{ p{3cm}llllllll}
			\hline
			Name & Dim & Pll & Com & Rflx & MFl & Turb & CGE & OS\\
			\hline
			DL\_MESO (LBE) & 2, 3 & MPI/OMP & Y & Y & Y & Y & Y & W, M, L\\
			ESPResSo & 1, 2, 3 & CUDA/MPI & Y & Y & Y & Y & Y & M, L\\
			ESPResSo++ & 1, 2, 3 & MPI & Y & Y & Y & Y & Y & L\\
			HemeLB & 3 & MPI & Y & Y & Y & Y & Y & L\\
			laboetie & 2, 3 & MPI & Y & Y & Y & Y & Y & L\\
			LatBo.jl & 2, 3 & - & Y & Y & Y & N & Y & L\\
			LB2D-Prime & 2 & MPI & Y & Y & Y & Y & Y & W, L\\
			LB3D & 3 & MPI & N & Y & Y & Y & Y & L\\
			LB3D-Prime & 3 & MPI & Y & Y & Y & Y & Y & W, L\\
			lbmpy & 2, 3 & CUDA & Y & Y & Y & Y & Y & L\\
			lettuce & 2, 3 & CUDA & Y & Y & Y & Y & Y & W, M, L\\
			LIMBES & 2 & OMP & Y & Y & N & N & Y & L\\
			Ludwig & 2, 3 & MPI & Y & Y & Y & Y & Y & L\\
			LUMA & 2, 3 & MPI & Y & Y & Y & Y & Y & W, M, L\\
			MechSys & 2, 3 & - & Y & Y & Y & Y & Y & L\\
			MP-LABS & 2, 3 & MPI/OMP & N & Y & Y & N & N & L\\
			OpenLB & 1, 2, 3 & MPI & Y & Y & Y & Y & Y & W, M, L\\
			Palabos & 2, 3 & MPI & Y & Y & Y & Y & Y & W, L\\
			pyLBM & 1, 2, 3 & MPI & Y & Y & N & Y & Y & W, M, L\\
			Sailfish & 2, 3 & CUDA & Y & Y & Y & Y & Y & M, L\\
			SunlightLB & 3 & - & Y & Y & N & N & Y & L\\
			TCLB & 2, 3 & CUDA/MPI & Y & Y & Y & Y & Y & L\\
			waLBerla & 2, 3 & MPI & Y & Y & Y & Y & Y & L\\
			\hline
		\end{tabular}
		\caption{Features of Software Packages (Dim for Dimension (1, 2, 3),
			Pll for Parallel (CUDA, MPI, OpenMP (OMP)), Com for Compressible (Yes or No), Rflx for
			Reflexive Boundary Condition (Yes or No), MFl for Multi-fluid (Yes or
			No), Turb for Turbulent (Yes or No), CGE for Complex Geometries (Yes or No), OS for Operating System (Windows (W), macOS (M), Linux (L)))} \label{tbl_features}
	\end{center}
\end{table}

@@

All of the 23 software packages that were tested have installation instructions.
As noted previously, many of the 23 software packages that were part of the
original long list of 45 packages were removed due to not including
documentation or installation instructions. Of the 23 software packages that
were tested, most had installation instructions located in one place, often in
an instruction manual or on a web-page. Sometimes, like with Ludwig, incomplete
installation instructions are found on a home page, with more detailed
instructions located on another web-page, or within the documentation.
Maintainability and correctness of these instructions could be improved if all
the instructions were in one location.

All 23 packages on the short list can be installed on some Unix-like systems.
Seven packages could be installed on Windows, and five on macOS. Operating
system compatibility is found in the documentation of 19 software packages. All
but one of the software packages (TCLB) were tested on Ubuntu for this state of
the practice assessment. TCLB was tested on CentOS, since this operating system
is mentioned in its installation instructions.

@@

Sixteen software packages require less than 10 dependencies to be installed. All
but one software package (LatBo.jl) require less than 20 dependencies. Some
packages may automatically install additional dependencies in the background.
Eighteen of the software packages do not explicitly indicate software dependency
versions. Some software package installation issues, specifically those
occurring when manual installation of dependencies is required, may be avoided
if versions of dependencies are specified. Fifteen software packages do not have
detailed instructions for installing dependencies. Sixteen software packages
have less than 10 manual installation steps. If dependencies are installed in
one command then none of the software packages take more than 20 steps to
install. The average number of steps is about eight, and the fewest is two
(LB3D-Prime).

@@

Figure~\ref{Fig_Installability} shows the installability ranking of the software
packages using AHP. Software packages with a higher score (MechSys, Palabos,
ESPResSo, OpenLB) tend to have one set of linear installation instructions that
are written as if the person doing the installation has none of the dependencies
installed. The instructions often list compatible operating system versions and
include instructions for the installation of dependencies. The top ranked
packages often incorporate some sort of automation of the installation process
and have fewer manual installation steps. The number of dependencies a package
has does not correlate with a higher score. The ability to validate the
installation process, often through tutorials or test examples that include
expected output, is correlated with a higher score. Furthermore, the top six
ranked packages are noted as being alive.

@@

Sixteen of the software packages include a requirements specification artifact
or explicitly reference domain theory, often only the latter. Software packages
that distribute requirements specification information, such as DL\_MESO (LBE),
generally keep it brief and include it within other documentation. This artifact
is often found within a user manual, on a web-page, or is mentioned in related
publications. In the latter case the user may need to spend significant time to
find this information.

Document generation tools are explicitly used by 12 software packages. Sphinx is
used by eight of them, and Doxygen is used by seven. Several of the packages use
both.

Tutorials are available for 18 of the software packages. Generally they are
linearly written and easy to follow. However, only eight tutorials provide an
expected output. It is not possible to verify the correctness of the output of
the software packages that are missing this key information. In these cases the
user may need to assume correctness if there are no visible errors.

Unit tests are only explicitly available for one of the software packages,
Ludwig. Code modularization of most packages allow for users to create tests
with varying degrees of effort. These tests allow developers and users to verify
the correctness of fragments of the source code, and in doing so better assess
the correctness of the entire package.

The use of continuous integration tools and techniques alludes to a more refined
development process where faults are isolated and better recognized. Only two of
the packages (ESPResSo, Ludwig) mentioned applying the practice of continuous
integration in their development process.

@@

Figure~\ref{Fig_CorrectnessVerifiability} shows the surface correctness and
verifiability ranking of the software packages using AHP. Software packages with
a higher score tend to have a visible requirements specification or references
to theory documentation. They also explicitly use at least one document
generation tool that builds confidence of correctness. The top ranked software
packages all include an easy to follow getting started tutorial, and most of
these include expected output. Only the top ranked package, Ludwig, provided
unit testing. It and the second ranked package, ESPResSo, explicitly
incorporated continuous integration in the development process. Furthermore,
eight of the top 10 ranked packages are noted as being alive.

@@

Of the 13 software packages that installed correctly and also have tutorials,
four (pyLBM, ESPResSo++, LIMBES, Ludwig) broke during tutorial testing. All of
these instances resulted in an error message being displayed. One error (pyLBM)
was due to a missing tutorial dependency, another (Ludwig) was due to an invalid
command despite following the tutorial, and the final two errors were vague
execution errors. Of the four broken tutorial instances, only the one that was
missing a dependency was recoverable.

@@

Figure~\ref{Fig_Reliability} shows the surface reliability ranking of the
software packages using AHP. Software packages with a high score either did not
break during installation, or the broken installation was recoverable. All of
the top five ranked packages have tutorials. One of these packages, pyLBM, broke
during tutorial testing, but a descriptive error message helped in recovery.
Furthermore, nine of the top 10 ranked packages are noted as being alive.

@@

Although the software packages all apply LBM to solve scientific computing
problems, the packages focus on varied CFD problems, with varying parameters,
and are technically different from each other (as shown in
Table~\ref{tbl_features}). Due to this, a comparison of performance is not
appropriate. In this project we instead looked through each software package's
artifacts for evidence that performance was considered. The artifacts of 17
software packages mentioned parallelization. This included GPU processing and
the CUDA parallel computing platform, which were mentioned in the artifacts of 6
packages (ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB). GPUs provide
superior processing power and speed compared to CPUs, and are often used for
scientific computing when a large amount of data is involved. The software
package TCLB is implemented in a highly efficient multi-GPU code to achieve
performance suitable for model optimization \citep{rutkowski2020open}. In the
Ludwig package, a so-called mixed mode approach is used where fine-grained
parallelism is implemented on the GPU, and MPI is used for even larger scale
parallelism \citep{gray2013ludwig}. While one software package (Sailfish)
required CUDA and GPU processing, some (ESPResSo, lbmpy, lettuce, pyLBM, TCLB)
have the option of using either the GPU or the CPU. The packages that require
GPU and CUDA have better performance at the expense of installability and
surface reliability.

@@

Software package artifacts were reviewed for the presence of a tutorial, a user
manual, documented user characteristics, and a user support model. In total 18
software packages have a tutorial, 13 have a user manual, and 11 have both. The
tutorials vary in scope and substance, and eight include expected output.
Most user manuals are in the form of a file that can be downloaded, while some
are rendered on a web-page. Some packages (waLBerla) do not have a user manual,
but do have useful documentation distributed on their web-pages. Expected user
characteristics are documented in four software packages (laboetie, LIMBES,
Ludwig, Palabos). Users are typically scientists or engineers. Their background
is often physics, chemistry, biophysics, or mathematics. All but one of the
packages (LIMBES) have a user support model, and many of them have multiple
avenues of user support. The most popular avenue of support is Git, followed by
email and forums. One software package (OpenLB) has an FAQ page.

@@

Issue tracking is used in 22 software packages, 15 of which use Git, six use
email, and one (SunlightLB) uses SourceForge. Most software packages that use
Git have most of their issues closed, and only three (laboetie, lettuce,
Sailfish) have less than 50 percent of their issues closed. Four of the top five
overall ranked packages (Ludwig, ESPResSo, Palabos, LUMA) have most of their
issues closed. Fourth ranked LUMA does not use Git. Alive packages (11 use Git
issue tracking) have 64\% of their issues closed, while dead packages (3 use Git
issue tracking) have 71\% of their issues closed. This information is presented
in Table~\ref{gitrepodata}. Furthermore, 13 packages that use Git for issue
tracking use GitHub as a version control system, while two (Palabos, waLBerla)
use GitLab. Of the other packages, one package (SunlightLB) uses CVS for issue
tracking, and seven packages do not appear to use any issue tracking system.

\begin{table}
	\begin{center}
		\begin{tabular}{ p{3.5cm}p{3.5cm}p{3.5cm}p{2.5cm} }
			\hline
			Name & $\%$ Issues Closed & $\%$ Code Comments & Status\\
			\hline
			DL\_MESO (LBE) & Not Git & 8.06& Alive\\
			ESPResSo & 89.26 & 21.78 & Alive\\
			ESPResSo++ & 66.28 & 17.10 & Alive\\
			HemeLB & No Issues & 16.68 & Dead\\
			laboetie & 18.75 & 2.47 & Dead\\		
			LatBo.jl & 93.33 & 0.40 & Dead\\
			LB2D-Prime & Not Git & 13.61 & Dead\\
			LB3D & Not Git & 13.76 & Dead\\
			LB3D-Prime & Not Git & 14.34 & Dead\\
			lbmpy& 58.33  & 2.03 & Alive\\
			lettuce & 33.33 & 8.19 & Alive\\
			LIMBES & Not Git & 17.39 & Dead\\
			Ludwig& 60.00 & 20.70 & Alive\\
			LUMA& 85.71   & 0.20 & Alive\\
			MechSys & Not Git & 15.11 & Alive\\
			MP-LABS & 100.00 & 26.67 & Dead\\
			OpenLB & Not Git & 22.43 & Alive\\
			Palabos & 89.47 & 17.76 & Alive\\
			pyLBM & 66.67& 16.12 & Alive\\
			Sailfish & 22.22 & 9.26 & Alive\\
			SunlightLB & Not Git & 17.67 & Dead\\
			TCLB & 60.32 & 6.02 & Alive\\
			waLBerla & 72.90 & 22.62 & Alive\\
			\hline
		\end{tabular}
		\caption{Git Repository Data} \label{gitrepodata}
	\end{center}
	\end{table}

Software package code files were further measured for the percentage of code
that is comments. The findings are presented in Table~\ref{gitrepodata}.
Packages with a higher percentage of comments were designated as more
maintainable. Comments represent more than 10 percent of code files in 15
packages, and the average percentage of code comments is about 14 percent. Four
of the top five overall ranked packages (Ludwig, ESPResSo, Palabos, OpenLB) have
more than the average. Fifth ranked LUMA has only 0.2 percent comments, the
fewest of any package. This package has the most lines of source code, with over
four million. The next largest package is ESPResSo++ with one million.

@@

Figure~\ref{Fig_Maintainability} shows the maintainability ranking of the
software packages using AHP. Software packages with a high score provide version
numbers on documents and source code releases, have an abundance of high quality
artifacts, and use an issue tracking tool and version control system. These
packages also appear to reasonably handle issue tracking, having most of their
issues closed. Their code files are well commented with more than 10 percent of
the code being comments. Furthermore, nine of the top 10 ranked packages are
noted as being alive. MP-LABS is noted as being dead.\\

@@

\begin{table}
	\begin{center}
		\begin{tabular}{ p{3.5cm}p{2cm}p{2.5cm}p{2cm}p{2.5cm} }
			\hline
			Name & Text Files & Binary Files & LOC & Avg. LOC / Text File\\
			\hline
			DL\_MESO (LBE) & 310 & 51 & 170223 & 549\\
			ESPResSo & 1309& 86 & 186700 & 143\\
			ESPResSo++ & 5328& 66 & 969196 & 182\\
			HemeLB & 1065& 48 & 95104 & 89\\
			laboetie & 133& 1 & 48403 & 364\\		
			LatBo.jl & 41& 0 & 42172 & 1029\\
			LB2D-Prime & 82& 19 & 54755 & 668\\
			LB3D & 99 & 76 & 39766 & 402\\
			LB3D-Prime & 2 3& 6 & 12944 & 563\\
			lbmpy& 201 & 28 & 46489 & 231\\
			lettuce & 62 & 0 & 5529 & 89\\
			LIMBES & 26 & 1 & 4872 & 187\\
			Ludwig & 859 & 32 & 109811 & 128\\
			LUMA & 312 & 19 & 4370670 & 14000\\
			MechSys & 324 & 3 & 85543 & 264\\
			MP-LABS & 307 & 3 & 43124 & 140\\
			OpenLB & 1104 & 5 & 209034 & 189\\
			Palabos & 1829 & 67 & 547623 & 299\\
			pyLBM & 258 & 85 & 32314 & 125\\
			Sailfish & 632 & 11 & 69398 & 110\\
			SunlightLB & 36 & 1 & 7646 & 212\\
			TCLB & 535 & 7 & 43226 & 81\\
			waLBerla & 2395 & 67 & 848146 & 353\\
			\hline
		\end{tabular}
		\caption{Module Data} \label{moduledata}
	\end{center}
\end{table}

@@

Figure~\ref{Fig_Understandability} shows the surface understandability ranking
of the software packages using AHP. Software packages with a high score have a
consistent indentation and formatting style, and consistent, distinctive, and
meaningful code identifiers. They also have symbolic constants, and explicitly
identify mathematical and LBM algorithms. Their comments are clear and indicate
what is being done in the source code. The source code is well modularized and
structured. Furthermore, four of the top five ranked packages are noted as being
alive.

@@

\subsection{What are the pain points?} \label{painpoints}
